Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers

### 2023-04-28 manuscript/manuscript.tex
changes

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 4020}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers},
  pdfauthor={Robert T. Thibault1,2, Marton Kovacs3,4, Tom E. Hardwicke5, Alexandra Sarafoglou6, John P. A. Ioannidis1,7, \& Marcus R. Munafò2,8},
  pdflang={en-EN},
  pdfkeywords={Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
\justifying\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis. An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW). In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it. The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset. ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance. The present survey aimed to garner information from the research community where ECAWs could be applied---employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example. \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC's transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}. We received 107 responses, for a 10\% response rate. The results suggest that---at least among our sample of respondents---ECAWs hold the potential to serve their intended purpose and appear relatively acceptable. For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55\% who agreed). However, as many as 25\% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44\% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions. \hfill \break \hfill \break \textbf{Preregistration:} \url{https://osf.io/g2fw5} Deviations from the preregistration are outlined in Supplementary Material A.
}

@@

Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2016; Collaboration, 2015; Errington et al., 2021; Ioannidis, 2005, 2008). These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices. These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see Prasad and Cifu (2019)).

@@

We call this process an Explore and Confirm Analysis Workflow (ECAW). Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly. ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset. Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., Bakker et al. (2020)), ECAWs provide an executable analysis script. Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration. ECAWs would not fully solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.

@@

One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (Sarafoglou, Hoogeveen, and Wagenmakers (2023)). The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded). Based on self-reports from the participants, the researchers found that two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.

Some disciplines, such as particle physics, implement blind data analyses regularly (Klein and Roodman (2005); MacCoun and Perlmutter (2015)). A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al. (2020)) and is currently being used for Registered Reports based on a large psychology dataset (Schmidt et al. (2023)).

@@

Respondents had published a median of 10 (IQR 2 to 26) studies using preexisting observational data (Supplementary Figure C1). They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}. 62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/100) reported being less concerned (Supplementary Figure C2).

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 1, top panel). At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using preexisting observational data (Figure 1, bottom panel).

@@

\caption{\textbf{Figure 1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}. The bar charts in the bottom panel excluded responses of \emph{``I don't understand the question''} (n = 3; 2).}\label{fig:typicallyEcawPlot}
\end{figure}

@@

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2). About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103). 77\% (79/103) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both exploratory (93\%; 96/103) and confirmatory (87\%; 90/103) analyses at least sometimes.

@@

26\% (26/101) of respondents agreed (versus 45\%; 45/101 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). 53\% (50/94) agreed (20\%; 19/94 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 55\% (53/96) agreed (10\%; 10/96 disagreed) that ALSPAC should run a study on ECAWs. 46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using typical preregistration.

@@

\caption{\textbf{Figure 3. Responses to survey questions about using ECAWs.} These bar charts exclude responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1; respectively from top to bottom), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question \protect\footnotemark[3]. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study'' \protect\footnotemark[4].}\label{fig:alspacPlot}
\end{figure}

@@

\textbf{\emph{Table 1. Recurring topics in responses to the open-ended survey questions.}} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of (92) responses from (55) unique respondents. A complete list of responses are viewable in the open data {[}LINK{]}. We synthesized the responses to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs. On the right side of the table, we provide our reflection on each topic.

@@

Results can be explored interactively by following the instructions for running the code at \url{https://github.com/RobbyTi/ECAW}. Based on the sample size of 103 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.

@@

The survey results suggest that a trial of ECAWs with ALSPAC could be feasible for at least three reasons. First, ECAWs are possible because most respondents reported performing confirmatory analyses with analysis scripts written using programming languages. Second, ECAWs are relevant because many respondents reported limited use of other methods to improve rigour and reproducibility---including preregistration, sharing analysis scripts, and blinding analysts. Moreover, although participants generally agreed that findings from observational research using pre-existing data are trustworthy and reproducible, they also believed that ECAWs would make research findings more trustworthy and more reproducible. Third, ECAWs appear relatively acceptable. For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 45\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.

@@

Given the response rate of 10\%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices. Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2). ECAWs may be less acceptable among non-responders and non-response may be a mark of lack of interest in the concept. Nonetheless, the absolute number of responders suggests that there is an audience of researchers who might be interested to pursue this approach.

@@

The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent. For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., Forscher et al. (2020); Schmidt et al. (2023)). Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data. For example, a researcher trying to answer a question about health and development with ALSPAC data, may also be able to answer that question with another cohort dataset. However, a researcher trying to answer a question about the population of a specific country may need access to that country's census data, regardless of the workflow required by that data management organization. A final consideration is that user numbers and engagement may increase if researchers feel that ECAWs increase the trustworthiness of their findings and others come to associate research from datasets using ECAWs as more open and rigorous.

@@

@@

@@

  \item We rephrased our second objective to be more accurate. The preregistration reads “2.    To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices (survey Blocks 3-6). This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability. ”. The manuscript reads “(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.”
  \item Marton Kovacs was added as a contributor during data collection. This led to a few of the projected contributor roles outlined in the preregistration to be different from the final contributor roles outlined in the manuscript.
  \item The preregistration stated that: “We will tabulate descriptive summary statistics for all the survey questions.” Instead of tabulating the results, the manuscript presents this data in figures. We feel that the figures are easier to digest as compared to tabulated data.
  \item The preregistration stated that: “we   will   present   results   that   include   responses   from participants who did not complete the entire survey, alongside the associated response rate for each question.”. The results from all participants are presented in Supplementary Material D. However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 107 times and partially completed 27 times, leading to a response rate of 10% for complete surveys and 12% for incomplete surveys.”
  \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
  \item We performed a brief visual inspection of the data presented in spreadsheet format. Three participants clicked through to the end of the survey—so they were coded as completing the survey—however, they only responded to a few of the first questions. One other participant appears to have provided non-sincere responses in that all their responses were neutral or skipped and they completed the survey in 146 seconds. We retained these four participants as partially completed responses (results in Supplementary Material D), but removed them from the main dataset. These decisions were not preregistered.
  \item We did not preregister which results we would present in the abstract. We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 25\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”

@@

\caption{\textbf{Figure C1.} Responses to survey question about number of relevant publications}\label{fig:numStudiesPlot}
\end{figure}

@@

@@

\caption{\textbf{Figure C2. Responses to survey question about number of relevant publications}}\label{fig:concernedPlot}
\end{figure}

@@

Respondents had published a median of 10 (IQR 2 to 25) studies using preexisting observational data. They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) \footnote[1]{Participants could select multiple responses to this survey question.}. 61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/101) reported being less concerned.

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Figure D2, top panel). At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 75/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using preexisting observational data (Figure 1, bottom panel).

@@

\caption{\textbf{Figure D1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}. The bar charts in the bottom panel excluded missing responses (n = 12; 12) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlotAll}
\end{figure}

@@

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Figure 2D). About half reported sharing their analysis scripts never or almost never (19\%; 23/123), or sometimes (29\%; 36/123). 69\% (85/123) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both exploratory (87\%; 107/123) and confirmatory (82\%; 101/123) analyses at least sometimes.

@@

25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3D). 51\% (50/98) agreed (21\%; 21/98 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 54\% (54/100) agreed (11\%; 11/100 disagreed) that ALSPAC should run a study on ECAWs. 45\% (45/100) agreed (22\%; 22/100 disagreed) that they would prefer using an ECAW than using typical preregistration.

@@

\protect\footnotetext[2]{Another four respondents agreed with at least the first and second question, which appear contradictory. We did not preregister these considerations. More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
\protect\footnotetext[3]{We intended for these questions to ask about ALSPAC running a trial on ECAWs. However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)\newline\indent Word count: 5036}
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers},
  pdfauthor={Robert T. Thibault1,2, Marton Kovacs3,4, Tom E. Hardwicke5, Alexandra Sarafoglou6, John P. A. Ioannidis1,7, \& Marcus R. Munafò2,8},
  pdflang={en-EN},
  pdfkeywords={Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
\justifying\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis. An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW). In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it. The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset. ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance. The present survey aimed to garner information from the research community where ECAWs could be applied---employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example. \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC's transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}. We received 103 responses, for a 9\% response rate. The results suggest that---at least among our sample of respondents---ECAWs hold the potential to serve their intended purpose and appear relatively acceptable. For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55\% who agreed). However, as many as 26\% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 45\% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions. \hfill \break \hfill \break \textbf{Preregistration:} \url{https://osf.io/g2fw5} Deviations from the preregistration are outlined in Supplementary Material A.
}

@@

Many published research findings are non-reproducible and potentially false or misleading (Camerer et al., 2016; Collaboration, 2015; Errington et al., 2021; Ioannidis, 2005, 2008). These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices. These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see Prasad \& Cifu, 2019).

@@

We call this process an Explore and Confirm Analysis Workflow (ECAW). Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly. ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset. Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., Bakker et al., 2020), ECAWs provide an executable analysis script. Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration. ECAWs would not fully solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.

@@

One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (Sarafoglou, Hoogeveen, \& Wagenmakers, 2023). The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded). Based on self-reports from the participants, the researchers found that two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.

Some disciplines, such as particle physics, implement blind data analyses regularly (Klein \& Roodman, 2005; MacCoun \& Perlmutter, 2015). A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (Forscher et al., 2020) and is currently being used for Registered Reports based on a large psychology dataset (Schmidt et al., 2023).

@@

Respondents had published a median of 10 (IQR 2 to 26) studies using preexisting observational data (Supplementary Figure C1). They reported using the following programming languages or software packages: R (\emph{n} = 65), Stata (\emph{n} = 48), SPSS (\emph{n} = 17), SAS (\emph{n} = 15), Python (\emph{n} = 6), Mplus (\emph{n} = 3), Bash (\emph{n} = 2), MATLAB (\emph{n} = 1), Nextflow (\emph{n} = 1), and plink2 (\emph{n} = 1) (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}. 62\% (62/100) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; 6\% (6/100) reported being less concerned (Supplementary Figure C2).

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (72\%; 74/103) and reproducible (79\%; 81/103) (Figure 1, top panel). At the same time, many believed that a study using an ECAW would be \emph{more} trustworthy (70\%; 70/100) and \emph{more} reproducible (68\%; 69/101) compared to a typical study using pre-existing observational data (Figure 1, bottom panel).

@@

\caption{\textbf{Figure 1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option \emph{``I don't understand the question''}. The bar charts in the bottom panel excluded responses of \emph{``I don't understand the question''} (\emph{n} = 3; 2; respectively from top to bottom).}\label{fig:typicallyEcawPlot}
\end{figure}

@@

Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (36\%; 37/103), or sometimes (25\%; 26/103) (Figure 2). About half reported sharing their analysis scripts never or almost never (20\%; 21/103), or sometimes (32\%; 33/103). 77\% (79/103) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both confirmatory (87\%; 90/103) and exploratory (93\%; 96/103) analyses at least sometimes.

@@

26\% (26/101) of respondents agreed (versus 45\%; 45/101 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Figure 3). 53\% (50/94) agreed (20\%; 19/94 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 55\% (53/96) agreed (10\%; 10/96 disagreed) that ALSPAC should run a study on ECAWs. Finally, 46\% (43/94) agreed (22\%; 21/94 disagreed) that they would prefer using an ECAW than using conventional preregistration.

@@

\caption{\textbf{Figure 3. Responses to survey questions about using ECAWs.} These bar charts exclude responses of \emph{``I don't understand the question''} (\emph{n} = 0; 4; 1; 1), and responses of \emph{``Unsure''} (\emph{n} = 2; 5; 6; 8). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word ``less'' in the first question \protect\footnotemark[3]. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term ``study'' \protect\footnotemark[4].}\label{fig:alspacPlot}
\end{figure}

@@

\textbf{\emph{Table 1. Recurring topics in responses to the open-ended survey questions.}} The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of 92 responses from 55 unique respondents. A complete list of responses are viewable in the open data {[}LINK{]}. We synthesized the responses to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs. On the right side of the table, we provide our reflection on each topic.

@@

Results can be explored interactively by running the code available at \url{https://github.com/RobbyTi/ECAW}. Based on the sample size of 103 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.

@@

The survey results suggest that a trial of ECAWs with ALSPAC could be feasible for at least three reasons. First, ECAWs are possible because most respondents reported performing confirmatory analyses with analysis scripts written using programming languages. Second, ECAWs are relevant because many respondents reported limited use of other methods to improve rigour and reproducibility---including preregistration, sharing analysis scripts, and blinding analysts. Moreover, although participants generally agreed that findings from observational research using pre-existing data are trustworthy and reproducible, they also believed that ECAWs would make research findings more trustworthy and more reproducible. Third, ECAWs appear relatively acceptable. For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, and 26\% agreed that they would be less likely to use the ALSPAC dataset if they were required to use an ECAW.

@@

Given the response rate of 9\%, our results represent the opinions of a select set of researchers who may be more interested and involved in reproducible research practices. Indeed, respondents themselves believed that they were more concerned about reproducibility than other researchers in their field (Supplementary Figure C2). ECAWs may be less acceptable among non-responders and non-response may be a mark of lack of interest in the concept. Nonetheless, the absolute number of responders suggests that there is an audience of researchers who might be interested to pursue this approach.

@@

The stakes are lower in cases where a static final dataset already exists and concerns about funding are absent. For example, researchers with an interest in rigorous analyses and who control access to a dataset have already employed ECAW-like workflows (e.g., Forscher et al., 2020; Schmidt et al., 2023). Concerns about a reduction in user numbers and engagement may also be less relevant for datasets containing unique data. For example, a researcher trying to answer a question about health and development with ALSPAC data, may also be able to answer that question with another cohort dataset. However, a researcher trying to answer a question about the population of a specific country may need access to that country's census data, regardless of the workflow required by that data management organization. A final consideration is that user numbers and engagement may increase if researchers feel that ECAWs increase the trustworthiness of their findings and others come to associate research from datasets using ECAWs as more open and rigorous.

@@

\leavevmode\vadjust pre{\hypertarget{ref-boyd_cohort_2013}{}}%
Boyd, A., Golding, J., Macleod, J., Lawlor, D. A., Fraser, A., Henderson, J., \ldots{} Davey Smith, G. (2013). Cohort profile: The {``children of the 90s''}---the index offspring of the {Avon} {Longitudinal} {Study} of {Parents} and {Children}. \emph{International Journal of Epidemiology}, \emph{42}(1), 111--127.

@@

\leavevmode\vadjust pre{\hypertarget{ref-fraser_cohort_2013}{}}%
Fraser, A., Macdonald-Wallis, C., Tilling, K., Boyd, A., Golding, J., Davey Smith, G., \ldots{} Ness, A. (2013). Cohort profile: The {Avon} {Longitudinal} {Study} of {Parents} and {Children}: {ALSPAC} mothers cohort. \emph{International Journal of Epidemiology}, \emph{42}(1), 97--110.

@@

  \item We rephrased our second objective to be more accurate. The preregistration reads “2.    To use these insights to inform future research on how data management organizations can encourage rigorous and reproducible research practices (survey Blocks 3-6). This objective includes assessing and refining potential interventions—such as ECAWs—and assessing their acceptability. ”. The manuscript reads “(2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.”
  \item Marton Kovacs was added as a contributor during data collection. This led to a few of the projected contributor roles outlined in the preregistration to be different from the final contributor roles outlined in the manuscript.
  \item The preregistration stated that: “We will tabulate descriptive summary statistics for all the survey questions.” Instead of tabulating the results, the manuscript presents this data in figures. We feel that the figures are easier to digest as compared to tabulated data.
  \item The preregistration stated that: “we   will   present   results   that   include   responses   from participants who did not complete the entire survey, alongside the associated response rate for each question.”. The results from all participants are presented in Supplementary Material D. However, instead of reporting the associated response rate for each question, we simply state that “The survey was completed 103 times and partially completed 20 times, leading to a response rate of 9% for complete surveys and 11% for at least partially complete surveys.”
  \item The preregistration and survey use the term “typical preregistration”. We changed this to “conventional preregistration” in the manuscript because we believe it is the more appropriate term.
  \item We performed a brief visual inspection of the data presented in spreadsheet format. Three participants clicked through to the end of the survey—so they were coded as completing the survey—however, they only responded to a few of the first questions. One other participant appears to have provided non-sincere responses in that all their responses were neutral or skipped and they completed the survey in 146 seconds. We retained these four participants as partially completed responses (results in Supplementary Material D), but removed them from the main dataset. These decisions were not preregistered.
  \item We did not preregister which results we would present in the abstract. We decided to report the results with the highest and lowest percentage about the acceptability of ECAWs: “For example, only 10\% of respondents disagreed that ALSPAC should run a study on ECAWs, but as many as 26\% of respondents agreed they would be less willing to use ALSPAC data if they were required to use an ECAW.”

@@

@@

\caption{\textbf{Figure C1. Responses to survey question about number of relevant publications.}}\label{fig:numStudiesPlot}
\end{figure}

@@

\caption{\textbf{Figure C2. Responses to survey question about number of relevant publications.} We excluded missing responses from this figure (\emph{n} = 3).}\label{fig:concernedPlot}
\end{figure}

@@

Respondents had published a median of 10 (IQR 2 to 25) studies using pre-existing observational data. They reported using the following programming languages or software packages: R (n = 65), Stata (n = 48), SPSS (n = 17), SAS (n = 15), Python (n = 6), Mplus (n = 3), Bash (n = 2), MATLAB (n = 1), Nextflow (n = 1), and plink2 (n = 1) \footnote[1]{Participants could select multiple responses to this survey question.}. 61\% (62/101) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses pre-existing observational data; 6\% (6/101) reported being less concerned.

@@

Most respondents agreed that studies that analyze pre-existing observational datasets are trustworthy (71\%; 87/123) and reproducible (77\%; 93/121) (Supplementary Figure D1, top panel). At the same time, many agreed that a study using an ECAW would be \emph{more} trustworthy (70\%; 75/107) and \emph{more} reproducible (69\%; 74/108) compared to a typical study using pre-existing observational data (Supplementary Figure D1, bottom panel).

@@

\caption{\textbf{Figure D1. Responses to the survey questions on trustworthiness and reproducibility of observational research with pre-existing data and ECAWs.} The survey defined trustworthy as ``meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust''. The survey defined reproducible ``in the sense that other researchers re-analysing the data with the same research question would produce similar results.'' For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel excluded missing responses (n = 0; 2) and had no selection of the option \emph{``I don't understand the question''}. The bar charts in the bottom panel excluded missing responses (n = 12; 12) and responses of \emph{``I don't understand the question''} (n = 4; 3).}\label{fig:typicallyEcawPlotAll}
\end{figure}

@@

Over half of respondents reported that their studies using pre-existing observational data are preregistered never or almost never (33\%; 41/123), or sometimes (21\%; 26/103) (Supplementary Figure 2D). About half reported sharing their analysis scripts never or almost never (19\%; 23/123), or sometimes (29\%; 36/123). 69\% (85/123) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both confirmatory (82\%; 101/123) and exploratory (87\%; 107/123) analyses at least sometimes.

@@

25\% (26/106) of respondents agreed (versus 44\%; 47/106 who disagreed) that they would be less willing to use ALSPAC data if they were required to use an ECAW (Supplementary Figure 3D). 51\% (50/98) agreed (21\%; 21/98 disagreed) that they would opt-in if ALSPAC ran a study on ECAWs. 54\% (54/100) agreed (11\%; 11/100 disagreed) that ALSPAC should run a study on ECAWs. 45\% (45/100) agreed (22\%; 22/100 disagreed) that they would prefer using an ECAW than using conventional preregistration.

@@
